---
title: "Bayesian deconvolution of delayed incidence, with random walk"
output: html_notebook
---

This notebook simulates a stochastic SEIR, convolves it with a discretized
gamma distribution, makes noisy observations, and infers the undelayed signal
using the correct delay distribution, assuming a random walk on
log(incidence).

The model assumes the following:

* The observations follow a Poisson distribution whose mean is equal to the
  convolution of the unobserved state with the delay distribution, plus a
  nonzero floor (to allow for zeros in the delay distribution)
* Log(unobserved states) follow a random walk (Brownian otion) whose standard
  deviation is inferred.

## Preamble

```{r}
library(ggplot2)
library(dplyr)
library(rstan)
source('../R/util.R')
source('../R/seir.R')
```


## Simulate stochastic SEIR

```{r}
sim_df <- {
  N <- 1e5
  E_frac <- 0.01
  E_init <- N * E_frac
  simulate_seir(
    arnaught = 2.0,
    t_E = 2,
    t_I = 4,
    N = N,
    S_init = N - E_init,
    E_init = E_init,
    I_init = 0,
    n_t = 50,
    n_steps_per_t = 10,
    method = 'stochastic'
  ) %>% filter(time > 0)
}

ggplot(sim_df, aes(x = time, y = dS)) +
  geom_col()
```

## Make a delay distribution from a discretized gamma

```{r}
delay_dist <- {
  t <- seq(0, 39)
  p_unnorm <- pgamma(t + 1, shape = 2.5, rate = 0.25) - pgamma(t, shape = 2.5, rate = 0.25)
  
  data.frame(delay = t, pmf = p_unnorm / sum(p_unnorm))
}

ggplot(delay_dist, aes(x = delay, y = pmf)) +
  geom_col()
  
```

## Delay incidence curve (dS) and make some right-censored, noisy observations

```{r}
p_obs <- 0.5
obs_df <- {
  y <- convolve_delay(sim_df$dS, delay_dist$pmf)[1:50]
  data.frame(time = 1:length(y), y = rpois(length(y), p_obs * y))
}

ggplot(obs_df, aes(x = time, y = y)) +
  geom_col()
```

## Deconvolve

```{r}
model <- stan_model('../stan/deconvolve-randomwalk.stan')
```

```{r}
model_input_data <- list(
  delay_min = delay_dist$delay[1],
  n_delay = length(delay_dist$pmf),
  pmf_delay = delay_dist$pmf,
  t_obs_min = obs_df$time[1],
  n_obs = nrow(obs_df),
  y_obs = obs_df$y,
  t_unobs_min = sim_df$time[1],
  n_unobs = nrow(sim_df),
  
  scale_x_unobs_init = max(obs_df$y),
  scale_sd_dlogx_unobs = max(log(obs_df$y) - lag(log(obs_df$y), 1), na.rm=TRUE)
)
```


## Fit

(Should just take a minute or so.)

```{r}
fit <- sampling(
  model, model_input_data, chains = 1, cores = 1,
  control = list(adapt_delta = 0.9)
)
params <- rstan::extract(fit)
```

## Plot the result

Legend:

* Dotted black line: observed data
* Solid black line: original signal multiplied by observation probability
* Solid red line: inferred signal
* Gray ribbon: 95% posterior interval

```{r}
result_df <- sim_df %>%
  left_join(
    tibble(
      time = 1:nrow(sim_df),
      xhat_mean = colMeans(params$x_unobs),
      xhat_025 = apply(params$x_unobs, 2, function(x) quantile(x, 0.025)),
      xhat_975 = apply(params$x_unobs, 2, function(x) quantile(x, 0.975))
    ),
    by = 'time'
  )

ggplot(result_df) +
  geom_ribbon(aes(x = time, ymin = xhat_025, ymax = xhat_975), fill = 'lightgray') +
  geom_line(aes(x = time, y = p_obs * dS)) +
  geom_line(aes(x = time, y = xhat_mean), color = 'red') +
  geom_line(data = obs_df, mapping = aes(x = time, y = y), lty = 2)
```

The estimates have much tighter confidence intervals, but perhaps a bit more
weirdness as you get into the censored region.
