---
title: "Richardson-Lucy-type deconvolution of delayed incidence"
output: html_notebook
---

NB: This code is NOT RECOMMENDED for use without extreme caution.

This notebook simulates a stochastic SEIR, convolves it with a discretized
gamma distribution, makes noisy observations, and infers the undelayed signal
using the correct delay distribution.

This notebook uses an implementation of the Richardson-Lucy-type deconvolution
used in [Goldstein et al. 2009 PNAS](https://doi.org/10.1073/pnas.0902958106).

The algorithm—or this implementation due to a bug—seems to be extremely sensitive to the initial guess.
Therefore, I don't recommend using it.

## Preamble

```{r preamble}
library(ggplot2)
library(dplyr)
library(rstan)
set.seed(1984)
```


## Simulate stochastic SEIR

```{r simulate stochastic SEIR}
source('../R/seir.R')
sim_df <- {
  N <- 1e5
  E_frac <- 0.01
  E_init <- N * E_frac
  simulate_seir(
    arnaught = 2.0,
    t_E = 2,
    t_I = 4,
    N = N,
    S_init = N - E_init,
    E_init = E_init,
    I_init = 0,
    n_t = 50,
    n_steps_per_t = 10,
    method = 'stochastic'
  ) %>% filter(time > 0)
}

ggplot(sim_df, aes(x = time, y = dS)) + geom_col()
```

## Make a delay distribution from a discretized gamma

```{r make delay distrib.}
delay_dist <- {
  t <- seq(0, 39)
  p_unnorm <- pgamma(t + 1, shape = 2.5, rate = 0.25) - 
    pgamma(t, shape = 2.5, rate = 0.25)
  data.frame(delay = t, pmf = p_unnorm / sum(p_unnorm))
}

ggplot(delay_dist, aes(x = delay, y = pmf)) + geom_col()
```

## Delay incidence curve (dS) and make some right-censored, noisy observations

```{r convolve}
source('../R/convolve_delay.R')
p_obs <- 0.5
obs_df <- {
  y <- convolve_delay(sim_df$dS, delay_dist$pmf)[1:50]
  data.frame(time = 1:length(y), y = rpois(length(y), p_obs * y))
}

ggplot(obs_df, aes(x = time, y = y)) +
  geom_col()
```

## Deconvolve

Despite inefficient implementation, this just takes a second:

```{r deconvolve}
source('../R/deconvolve.R')
result <- deconvolve_rltype_goldsteinetal(
  t_obs_min = obs_df$time[1],
  y_obs = obs_df$y,
  delay_min = delay_dist$delay[1],
  pmf_delay = delay_dist$pmf,
  t_unobs_min = sim_df$time[1],
  n_unobs = nrow(sim_df),
  n_iterations_max = 100,
  n_iterations = NULL
)
result$chi_square
result$n_iterations
```

## Plot the result

Legend:

* Dotted black line: observed data
* Solid black line: original signal multiplied by observation probability
* Solid red line: inferred signal

```{r plot}
result_df <- sim_df %>%
  left_join(
    tibble(
      time = 1:nrow(sim_df),
      xhat = result$x_unobs
    ),
    by = 'time'
  )

ggplot(result_df) +
  geom_line(aes(x = time, y = p_obs * dS)) +
  geom_line(aes(x = time, y = xhat), color = 'red') +
  geom_line(data = obs_df, mapping = aes(x = time, y = y), lty = 2)
```

This looks pretty good, but the values in the right-censored region are very unstable in terms of initial conditions.

In fact, everything is very unstable in terms of initial conditions.
Redoing everything with initial conditions bootstrap-sampled from the observed time series gives a strange result:

```{r strange things are happening to me}
{
  source('../R/deconvolve.R')
  result <- deconvolve_rltype_goldsteinetal(
    t_obs_min = obs_df$time[1],
    y_obs = obs_df$y,
    delay_min = delay_dist$delay[1],
    pmf_delay = delay_dist$pmf,
    t_unobs_min = sim_df$time[1],
    n_unobs = nrow(sim_df),
    n_iterations_max = 100,
    n_iterations = NULL,
    x_unobs_init = sample(obs_df$y, nrow(sim_df), replace = TRUE)
  )
  result$chi_square
  result$n_iterations
  
  result_df <- sim_df %>%
    left_join(
      tibble(
        time = 1:nrow(sim_df),
        xhat = result$x_unobs
      ),
      by = 'time'
    )
  
  ggplot(result_df) +
    geom_line(aes(x = time, y = p_obs * dS)) +
    geom_line(aes(x = time, y = xhat), color = 'red') +
    geom_line(data = obs_df, mapping = aes(x = time, y = y), lty = 2)
}
```
