---
title: "Richardson-Lucy Blind Deconvolution: In progress"
output: html_notebook
---

RL deconvolution analysis by Lauren McGough, Simulations adapted from Katie Gostic

Last updated: 6-16-2020

This notebook runs, but the "blind" part of the blind deconvolution isn't working yet.

Using blind deconvolution to compensate for a misspecified delay distribution. 

So far, find that the best parameters are: 
* num_blind_it_param <- 12 to 15
* num_delay_it_param <- 3
* num_u_it_param <- 3
but this is not universal among delay distributions.

Note that the mean of the delay distribution is reproduced pretty well, though its variance still isn't, and unfortunately, the delay distribution is so far pretty sensitive to the number of iterations.



```{r}
rm(list = ls())
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(cowplot)
library(EpiEstim)
library(fs)
library(magic)
library(corpcor)
theme_set(theme_bw())

#keeping delay distribution fixed this time
delay_distr_shape_mult_param <- 1.1 #going to assume a delay distribution with shape that is delay_distr_shape_mult*true_shape, scale is delay_distr_scale_mult*true_scale
delay_distr_scale_mult_param <- 1

num_blind_it_param <- 12
num_delay_it_param <- 3
num_u_it_param <- 3

intervention_length_name <- sprintf("_blind-it_%s_delay-it_%s_u-it_%s_shape_mult_%s_scale_mult_%s_1", toString(num_blind_it_param), toString(num_delay_it_param), toString(num_u_it_param), toString(delay_distr_shape_mult_param), toString(delay_distr_scale_mult_param))
intervention_file_name <- sprintf("intervention%s", intervention_length_name)
if (!dir.exists(intervention_file_name)) {
  dir.create(intervention_file_name)
}
```


##Richardson-Lucy method for deconvolving delay distribution and observations of cases

```{r}
if (file.exists(sprintf("%s/params.Rds", intervention_file_name))) {
  parlist <- readRDS(sprintf("%s/params.Rds", intervention_file_name))
} else {
  ## Set parameters to simulate SEIR epidemic
  parlist <- {
    list(
      N = 2e6, #total population size
      E_init = 0,
      I_init = 10,
      t_E = 4, # mean time in E (latent period)
      t_I = 4, # mean time in I (duration of infectiousness)
      n_t = 300, # total timesteps
      pre_intervention_R0 = 2, # Initial R0 before interventions
      intervention_R0 = 0.8, # Final R0 after interventions
      partially_lifeted_R0 = 0.9,
      intervention_time_1 = 62, # Timepoint at which intervention starts (at which underlying transmission rate begins to fall)
      intervention_time_2 = 62+200,
      days_intervention_to_min = c(7), # Days from intervention start until transmission rate hits min_R0
      days_to_Rt_rise = 1,
      model_types = c('seir'), # Can also choose sir
      methods = c('stochastic'), # could also choose ode, but would have to modify it to have integer case counts
      obs_1_offset = 10, #the first observation is at 1+length(delay dist)+obs_1_offset
      obs_2_offset = 75, #the last observation is at 1+length(delay dist)+obs_1_offset+obs_2_offset
      max_RL_it = 10,
      delay_distr_shape_mult = delay_distr_shape_mult_param,
      delay_distr_scale_mult = delay_distr_scale_mult_param,
      num_blind_it = num_blind_it_param,
      num_delay_it = num_delay_it_param,
      num_u_it = num_u_it_param
    )
  }  
  ## Derive the mean and variance of the serial interval from the input parameters
  parlist$true_mean_SI = (parlist$t_E+parlist$t_I)
  parlist$true_var_SI = 2*parlist$true_mean_SI^2
  saveRDS(parlist, file = sprintf("%s/params.Rds", intervention_file_name))
}


```


## Write function to estimate $R_t$ - Cori method
```{r}
## Output cori estimate with mean, CI and times given an input df, and the name of the incidence column
# Note that here, we're using the option that samples over serial intervals and std of serial intervals
get_cori <- function(df.in, 
                     icol_name, 
                     out_name = 'Cori',
                     window = 1, 
                     SI_mean=parlist$true_mean_SI, 
                     SI_var=2*(parlist$true_mean_SI/2)^2,
                     wend = TRUE){
  idat <- df.in %>%
    filter(get(icol_name) > 0 & !is.na(get(icol_name))) %>%
    complete(time = 2:max(time))%>%
    mutate_all(.funs = function(xx){ifelse(is.na(xx), 0, xx)})
  
  ts <- idat$time
  ts <- ts[ts > 1 & ts < (max(ts)-window+1)]
  te<- ts+(window-1)
  
  estimate_R(
    incid = pull(idat, eval(icol_name)),
    method = "parametric",
    config = make_config(
      list(
        mean_si = SI_mean,
        #min_mean_si = SI_mean -1,
        #max_mean_si = SI_mean + 1,
        #std_mean_si = 1.5,
        #std_std_si = 1.5,
        std_si = sqrt(SI_var),
        #min_std_si = sqrt(SI_var)*.8,
        #max_std_si = sqrt(SI_var)*1.2,
        #n1 = 50,
        #n2 = 100, 
        t_start=ts,
        t_end=te
        )
      )
    ) -> outs
  
  outs$R %>%
    mutate(time = if(wend == TRUE) t_end else ceiling((t_end+t_start)/2) ) %>%
    select(time, `Mean(R)`, `Quantile.0.025(R)`, `Quantile.0.975(R)`) %>%
    setNames(c('time', paste0(out_name, '.mean'), paste0(out_name, '.025'), paste0(out_name, '.975')))
}
```

## Simulate SEIR data

```{r}
## Simulate SEIR data using a stochastic (ode) model. Putting the R0 and simplots folders in the intervention file.
source('funs_simulation-sweep.R')
sim_sweep(parlist)
testplots(parlist)

file_move("R0-2.0", intervention_file_name)
file_move("simplots", intervention_file_name) 
```

## Extract SEIR simulation and plot.

- Incidence here is observed at time of infection (S -> E).
- SEIR simulation is stochastic.

```{r}
## Write a function to extract the simulation results as a data frame
stoch_df <- function(){
  readRDS(sprintf('%s/R0-%.1f/seir_%s_dec%.0f-%.0f_sim.rds',
                  intervention_file_name,
                  parlist$pre_intervention_R0, 
                  parlist$methods,
                  parlist$intervention_time_1, 
                  parlist$days_intervention_to_min))$sim_df 
}

stoch_df() %>%
ggplot() +
  geom_line(aes(x = time, y = incidence))+
  geom_vline(aes(xintercept = parlist$intervention_time_1), lty = 2)+ ## Dashed line where Rt starts to decrease
    geom_vline(aes(xintercept = parlist$intervention_time_2), lty = 2)+ ## Dashed line where Rt starts to decrease
  ggtitle('Daily infections, SEIR simulation') -> inc

stoch_df() %>% 
  ggplot()+
  geom_line(aes(x = time, y = true_r0)) +
  geom_hline(aes(yintercept = 1), lty = 2)+
  ylab('R0')+
  ggtitle('Input Rt values') -> R0

plot_grid(R0, inc, rel_heights = c(1,2), align = 'hv', nrow = 2)

stoch_df
```
## Forward step: Impute times of observation from SEIR times of infection 

To test, setting the observation delay to always be 1 day.

```{r}
source('funs_impute_obs_times.R')

## Set the delay distribution to observation
obs_delay_dist <- function(nn){
  r_inc_dist <- function(n){rgamma(n, shape = 5.8, scale = 0.95)} # Incubation period (infection -> symptoms)
  r_sym_to_obs_dist <- function(n){runif(n, 0, 4)} # Additional delay from symptoms -> observation
  r_inc_dist(nn) + r_sym_to_obs_dist(nn)
#  rep(20, nn)
}

misspec_obs_delay_dist <- function(nn){
  r_inc_dist <- function(n){rgamma(n, shape = 5.8*parlist$delay_distr_shape_mult, scale = 0.95*parlist$delay_distr_scale_mult)} # Incubation period (infection -> symptoms)
  r_sym_to_obs_dist <- function(n){runif(n, 0, 4)} # Additional delay from symptoms -> observation
  r_inc_dist(nn) + r_sym_to_obs_dist(nn)
#  rep(20, nn)
}

## Append number of new daily infections to simulation dataframe
sim_df <- stoch_df() %>%
  filter(time < max(time)) %>%
  mutate(
    new_infected = ifelse(is.na(dS), 0, dS))
```

## Define observation step
```{r, echo=FALSE}
#get_tObs_from_tInf(sim_df$new_infected, sim_df$time, obs_delay_dist, return_times = T)
sim_df %>%
  merge(
    get_tObs_from_tInf(sim_df$new_infected, sim_df$time, obs_delay_dist, return_times = T),
    by = 'time', all = TRUE) %>% 
  rename(new_observed = n) %>%
  as.tbl() -> sim_df

```

### Deconvolving using Richardson-Lucy method 

Plotting the delay distribution. *Assuming a misspecified delay distribution here.*
  
```{r, echo=FALSE}
#THE CORRECT DELAY DISTRIBUTION
obs_delay_dist(10000000)  %>%
#    ceiling() %>%
  tabulate()*1/10000000 -> true_delay_distr_vec #could also set nbins = const, which would set the length of the delay distribution to always be const, at the expense of potentially having 0's

#ASSUMING A MISSPECIFIED DELAY DISTRIBUTION
misspec_obs_delay_dist(10000000)  %>%
#    ceiling() %>%
  tabulate()*1/10000000 -> delay_distr_vec #could also set nbins = const, which would set the length of the delay distribution to always be const, at the expense of potentially having 0's

len_diff <- length(sim_df$new_infected)-length(delay_distr_vec)

#length(delay_distr_vec)

#plot(1:length(delay_distr_vec), delay_distr_vec)

```

```{r}

#delay_distr_vec <- c(delay_distr_vec, replicate(len_diff,0))
#length(delay_distr_vec)

#delay_distr_vec
```

```{r}

new_inf_no_na <- sim_df$new_infected #number of cases at the S -> E transition. This is what RL will be trying to reproduce.
#new_obs_no_na <- ifelse(is.na(sim_df$new_observed), 0, sim_df$new_observed)
#obs_cases <- get_tObs_from_tInf(new_inf_no_na, 
#                              times=1:length(new_inf_no_na), 
#                               r_delay_dist=obs_delay_dist,
#                               return_times = FALSE)
new_obs_no_na <- sim_df$new_observed #this is the "data" vector we will apply RL to. 


```

Doing a set of RL iterations.

```{r}

get_chi_sq <- function (vec1, vec2) {
  n0 <- length(vec1)
  (1/n0)*sum(((vec1 - vec2)^2/vec1))
}

do_RL <- function (init_guess, obs_dat, ker_vec, obs_1, obs_2, max_it){

  #d_obs_rescaled <- d_obs * q_j[(length(p)):length(u_obs_guess_rescaled)]

  #u_obs_rescaled <- do_RL(d_obs_rescaled, u_obs_guess_rescaled, length(p), p_ij_obs_rescaled, parlist$max_RL_it) #not sure this is quite right - do you change u   and p both on every iteration? - no, do a bunch of u iterations, then a bunch of p iterations, then a bunch of u iterations, etc.

  #u_obs_new <- u_obs_rescaled/q_j
  
  
  ker_mat_unnorm <- get_ker_ij(ker_vec, obs_1, obs_2)
  q_j <- colSums(ker_mat_unnorm)
  
  dim_ker_mat <- dim(ker_mat_unnorm)
  ker_mat <-  ker_mat_unnorm / matrix(q_j,nrow=dim_ker_mat[1],ncol=dim_ker_mat[2],byrow=TRUE)
  #u_obs_rescaled <- u_obs_guess_rescaled
  
  len_ker <- length(ker_vec)
  curr_guess <- init_guess * q_j
  
  obs_dat_re <- obs_dat * q_j[(length(q_j) - length(obs_dat) + 1):length(q_j)]
  #print(obs_dat_re)
  
  
  #chi_sq <- get_chi_sq(d_obs_rescaled, u_obs_rescaled[length(p):length(u_obs_rescaled)])
  chi_sq <- get_chi_sq(obs_dat_re, curr_guess[len_ker:length(curr_guess)])
  #u_obs_rescaled
  #chi_sq
  ind <- 1
  while (chi_sq > 1 & ind < max_it) {
    #c_obs <-p_ij_obs_rescaled %*% u_obs_rescaled
    
    obs_guess <- ker_mat %*% curr_guess
    #print(c_obs)
    
    #new_kernel_obs <- d_obs/c_obs
    new_ker <- obs_dat_re/obs_guess
    #print(t(new_kernel_obs)%*%p_ij_obs_rescaled)
    
    #print(u_obs_rescaled)
    #new_u_obs_rescaled <- u_obs_rescaled * t(t(new_kernel_obs) %*% p_ij_obs_rescaled)
    new_guess <- curr_guess * t(t(new_ker) %*% ker_mat)
    #print(new_u)
    
    #u_obs_rescaled <- new_u_obs_rescaled
    curr_guess <- new_guess
    
    chi_sq <- get_chi_sq(obs_dat_re, curr_guess[len_ker:length(curr_guess)])
    ind <- ind+1
  }
  curr_guess/q_j
}

get_ker_ij <- function (this_ker, obs_time_1, obs_time_2){
  ker_mat <- circulant(c(this_ker, rep(0, obs_time_2)))
  ker_ij_obs <- ker_mat[(obs_time_2-length(this_ker)+1):obs_time_2, obs_time_1:obs_time_2]
  #p_ij_obs <- pmat[obs_1:obs_2, 1:length(u_obs_guess)]
  ker_ij_obs <- t(ker_ij_obs)
}

```


Carrying out blind RL on observed section of the data.
```{r}
p_unnorm <- c(0.000001, ifelse(delay_distr_vec==0, 0.000001, delay_distr_vec)) #the first element is the probability of delay = 0
p <- p_unnorm/sum(p_unnorm)

d <- new_obs_no_na
obs_1 <- length(p)+1+parlist$obs_1_offset
obs_2 <- obs_1 + parlist$obs_2_offset
dat <- d[obs_1:obs_2]

u_guess <- c(dat, rep(max(dat[length(dat)],1), length(p)-1))
p_guess <- c(p, rep(0.000001, length(dat)-1)) #same length as u_guess

this_u <- u_guess
this_norm_u <- this_u/sum(this_u)
this_delay <- p_guess
this_norm_delay <- this_delay/sum(this_delay)

num_delay_it <- parlist$num_delay_it
num_u_it <- parlist$num_u_it

#blind deconvolution works by iterating RL, first trying to infer the correct delay distribution using the guess, 
#then trying to infer the correct incidence curve using the current guess for the delay distribution 
#issue: although when doing this once, we extend u_guess to take care of boundary conditions, in this case, we don't want to keep extending u and delay indefinitely
for (k in 1:parlist$num_blind_it) { #try always making the kernel be normalized
  new_u <- do_RL(this_u, dat, this_norm_delay, obs_1, obs_2, num_u_it)
  this_u <- new_u
  this_norm_u <- this_u/sum(this_u)
  new_delay <- do_RL(this_delay, dat, this_norm_u, obs_1, obs_2, num_delay_it) #this_u is a vector - do_RL needs to convert it to a matrix
  this_delay <- new_delay #NOT NORMALIZED
  this_norm_delay <- this_delay/sum(this_delay)
}

p <- this_norm_delay
u_obs_new <- this_u

fixed_delay_distr_unnorm <- c(0.000001, ifelse(delay_distr_vec==0, 0.000001, delay_distr_vec))
fixed_delay_distr_norm <- fixed_delay_distr_unnorm/sum(fixed_delay_distr_unnorm)
fixed_true_delay_distr_unnorm <- c(0.000001, ifelse(true_delay_distr_vec==0, 0.000001, true_delay_distr_vec)) #the first element is the probability of delay = 0
fixed_true_delay_distr_norm <- fixed_true_delay_distr_unnorm/sum(fixed_true_delay_distr_unnorm)

delay_df <- data.frame(1:length(fixed_true_delay_distr_norm), this_norm_delay[1:length(fixed_true_delay_distr_norm)], fixed_true_delay_distr_norm)
names(delay_df) <- c("time", "calculated_delay", "true_delay")
misspec_df <- data.frame("time" = 1:length(fixed_delay_distr_norm), "misspecified" = fixed_delay_distr_norm)

delay_df_merge <- merge(delay_df, misspec_df, "time")

```

```{r}
delay_df %>%
  pivot_longer(names_to = "delays", cols = c("calculated_delay", "true_delay")) %>%
  ggplot() +
  geom_line(aes(x = time, y = value, color = delays)) +
  xlab("time (days)") + 
  ylab("density") + 
  ggtitle("Delay distribution calculated through blind deconvolution") -> delay_df_plot

delay_df_plot

ggsave(sprintf("%s/delays_calculated_true.png", intervention_file_name))



delay_df_merge %>%
  pivot_longer(names_to = "delays", cols = c("calculated_delay", "true_delay", "misspecified")) %>%
  ggplot() +
  geom_line(aes(x = time, y = value, color = delays)) + 
  xlab("time (days)") + 
  ylab("density") + 
  ggtitle("Delay distribution calculated through blind deconvolution") -> delay_df_merge_plot

delay_df_merge_plot

ggsave(sprintf("%s/delays_calculated_true_misspecified.png", intervention_file_name))
```

  
```{r}
#The below are all equal: must keep track of the fact that there's an extra 0 in the front since this is 0-indexed
#plot(this_norm_delay[1:length(true_delay_distr_vec)])
#points(c(0.0000001, true_delay_distr_vec), col = "blue")
#points(c(0.0000001, delay_distr_vec), col = "red")
```

```{r}



length(u_obs_new)
length(c(obs_1 - length(fixed_delay_distr_norm)+1):obs_2)

inferred_df <- data.frame(c(obs_1 - length(fixed_delay_distr_norm)+1):obs_2, u_obs_new)
names(inferred_df) <- c('time', "new_inferred")
#print(inferred_df)

#TRY REDEFINING THE TIME SHIFTS HERE##############################
inferred_df$time <- inferred_df$time - 1
sim_df$time <- sim_df$time + 1

sim_df %>%
  merge(inferred_df, by = 'time', all = 'TRUE') %>%
  as.tbl()  %>%
  pivot_longer(-time, names_to = "data_type", values_to = "count") %>% 
  filter(data_type == "new_inferred"| data_type == "new_infected" | data_type == "new_observed") %>%
  ggplot() +
  geom_line(aes(x = time, y = count, color = data_type, linetype = data_type)) + 
  geom_vline(aes(xintercept = parlist$intervention_time_1), linetype = "dotted") +
  scale_linetype_manual(values=c("solid", "solid", "dotted")) + 
  scale_color_manual(values = c("blue", "red", "green")) +
  labs(color = "Infections", linetype = "Infections") + 
  ylab("count")+
  xlab("time (days)") + 
  ggtitle('Number of inferred cases from Richardson-Lucy') -> inferred_plot

inferred_plot

ggsave(sprintf("%s/all-case-curves.png", intervention_file_name))

length(fixed_delay_distr_norm)

```

The red lines mark the beginning and end of where we have data.

```{r}
# 
# 
# diff <-  u_obs_new - u_true[(obs_1 - length(p) + 1):obs_2]
# ggplot()+
#   geom_line(aes(x=(c(obs_1 - length(p) + 1):obs_2), diff, colour = 'red')) +
#   ylab("difference in number of infections")+
#   xlab("time (days)") + 
# #  geom_vline(aes(xintercept = obs_1, colour = 'red'), linetype = "dashed") + 
# #  geom_vline(aes(xintercept = obs_2, colour = 'red'), linetype = "dashed") + 
#   geom_hline(aes(yintercept = 0, colour = 'black'), linetype = "dashed")+
#   geom_vline(aes(xintercept = parlist$intervention_time_1, color = "blue"), linetype = "dashed") +
#   scale_color_discrete(name = "Legend", labels = c("Difference = 0", "Time of intervention", "Infections, inferred minus true")) +
#   ggtitle('Difference, inferred infections minus true infections, Richardson-Lucy')
# 
# ggsave(sprintf("%s/difference-cases-RL.png", intervention_file_name))
```

```{r}

u_true <- new_inf_no_na 
diff <-  u_obs_new - u_true[(obs_1 - length(fixed_delay_distr_norm) + 1):obs_2]

#diff <-  u_obs_new - u_true[(obs_1 - length(delay_distr_vec)):(obs_2-1)]
length(diff)
length((obs_1 - length(delay_distr_vec)):obs_2)

sim_df2 <- sim_df
diff_df <- data.frame(c((obs_1 - length(delay_distr_vec)):obs_2), diff)
diff_df
names(diff_df) <- c('time', "diff")

sim_df2 %>%
  merge(diff_df, by = 'time', all = 'TRUE') %>%
  as.tbl()  %>%
  pivot_longer(-time, names_to = "data_type", values_to = "count") %>%
  filter(data_type == "diff" & !is.na(count)) %>%
  ggplot() +
  geom_line(aes(x = time, y = count, color = data_type, linetype = data_type)) + 
  geom_hline(aes(yintercept = 0, color = 'zero', linetype = 'zero'))+
  geom_vline(aes(xintercept = parlist$intervention_time_1, color = 'intervention', linetype = 'intervention')) +
  scale_linetype_manual(values=c("solid", "dotted", "dashed")) + 
  scale_color_manual(values = c("blue", "black", "black")) +
 # scale_color_discrete(name = "Legend", labels = c("Difference = 0", "Time of intervention", "Infections, inferred minus true")) +
  labs(color = "Legend", linetype = "Legend") +
  ylab("count (difference)")+
  xlab("time (days)") + 
  ggtitle('Difference in number of inferred cases from Richardson-Lucy') -> diff_plot

diff_plot

ggsave(sprintf("%s/difference-cases-RL.png", intervention_file_name))

```
```{r}


cori_df <- get_cori(df.in = sim_df, icol_name = "incidence")
cori_inferred <- get_cori(df.in = inferred_df, icol_name = "new_inferred")

# cori_df %>%
#   filter(!is.na(Cori.mean)) %>%
#   ggplot() + 
#   geom_line(aes(x = time, y = Cori.mean)) + 
#   geom_ribbon(aes(x = time, ymin=Cori.025, ymax = Cori.975), alpha=0.3)
# 
# cori_inferred %>%
#   filter(!is.na(Cori.mean) & Cori.mean < 5) %>%
#   ggplot() + 
#   geom_line(aes(x = time, y = Cori.mean)) + 
#   geom_ribbon(aes(x = time, ymin=Cori.025, ymax = Cori.975), alpha=0.3)

all_cori_df <- merge(cori_df, cori_inferred, by = "time")
r0_df <- stoch_df()
merge(all_cori_df, r0_df, by = "time") -> new_all_df

new_all_df %>%
  filter(!is.na(Cori.mean.x) & !is.na(Cori.mean.y) & Cori.mean.y < 5 & Cori.975.y < 8) %>%
  ggplot() + 
  geom_hline(aes(yintercept = 1, linetype = "one", color = "one")) + 
  geom_line(aes(x = time, y = true_r0, color = "true_Rt", linetype = "true_Rt"), size = 1) +
  geom_line(aes(x = time, y = Cori.mean.x, color = "true_infections", linetype = "true_infections")) + 
  geom_line(aes(x = time, y = Cori.mean.y, color = "inferred_infections", linetype = "inferred_infections")) +
  geom_ribbon(aes(x = time, ymin=Cori.025.x, ymax = Cori.975.x, fill = "true_infections"), alpha=0.3, show.legend = FALSE) + 
  geom_ribbon(aes(x = time, ymin=Cori.025.y, ymax = Cori.975.y, fill = "inferred_infections"), alpha=0.3, show.legend = FALSE) +
  geom_vline(aes(xintercept = parlist$intervention_time_1, color = 'intervention', linetype = 'intervention')) + 
  scale_linetype_manual(values=c("solid", "dotted", "dashed", "solid", "solid")) + 
  scale_color_manual(values = c("red", "black", "black", "blue", "black")) +
  scale_fill_manual(values = c("red", "blue")) +
 # scale_color_discrete(name = "Legend", labels = c("Difference = 0", "Time of intervention", "Infections, inferred minus true")) +
  labs(color = "Legend", linetype = "Legend", fill = "Legend") +
  xlab("time (days)") + 
  ylab("Rt") +
  ggtitle('Calculations of Rt') -> new_all_plot

new_all_plot

ggsave(sprintf("%s/rt_plot.png", intervention_file_name))
```


```{r}

plot_grid(inferred_plot, diff_plot, new_all_plot, labels = "AUTO", ncol = 1, align = 'v') -> all_plot

all_plot

ggsave(sprintf("%s/all_plot.png", intervention_file_name))

```
```{r}
cori_x_no_na <- new_all_df$Cori.mean.x[!is.na(new_all_df$Cori.mean.x)]
fft_x <- abs(fft(cori_x_no_na))
plot(abs(fft(cori_x_no_na))/fft_x[1])
#print(new_all_df$Cori.mean.x)

cori_y_no_na <- new_all_df$Cori.mean.y[!is.na(new_all_df$Cori.mean.y)]
fft_y <- abs(fft(cori_y_no_na))
points(abs(fft(cori_y_no_na))/fft_y[1], col = "red")
#print(new_all_df$Cori.mean.y)
```